My Final College Paper

### 2021-04-24 writing /thesis_draft/index.Rmd
starting get my act together to format this thesis

---
title: 'My Final College Paper'
author: 'Margaret A. Slein'
date: 'May 2021'
institution: 'Reed College'
division: 'Mathematics and Natural Sciences'
advisor: 'Sam Fey'
# If you have more two advisors, un-silence line 7
#altadvisor: 'Your Other Advisor'
department: 'Mathematics'
degree: 'Bachelor of Arts'
knit: bookdown::render_book
site: bookdown::bookdown_site

# This will automatically install the {remotes} package and {thesisdown}
# Change this to FALSE if you'd like to install them manually on your own.
params:
  'Install needed packages for {thesisdown}': True

# Remove the hashtag to specify which version of output you would like.
# Can only choose one at a time.
output:
  thesisdown::thesis_pdf: default 
#  thesisdown::thesis_gitbook: default         
#  thesisdown::thesis_word: default
#  thesisdown::thesis_epub: default

# If you are creating a PDF you'll need to write your preliminary content 
# (e.g., abstract, acknowledgements) below or use code similar to line 25-26 
# for the .RMD files. If you are NOT producing a PDF, delete or silence
# lines 25-39 in this YAML header.
abstract: '`r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00-abstract.Rmd")), collapse = "\n  ")`'
# If you'd rather include the preliminary content in files instead of inline
# like below, use a command like that for the abstract above.  Note that a tab 
# is needed on the line after the `|`.
acknowledgements: |
  The future is bright, the future is dark

@@

dedication: |
  You can have a dedication here if you wish. 
preface: |
  This is an example of a thesis setup to use the reed thesis document class 
  (for LaTeX) and the R bookdown package, in general.

# Specify the location of the bibliography below
bibliography: bib/thesis.bib
# Download your specific csl file and refer to it in the line below.
csl: csl/apa.csl
lot: true
lof: true
---

@@

Welcome to the _R Markdown_ thesis template. This template is based on (and in many places copied directly from) the Reed College LaTeX template, but hopefully it will provide a nicer interface for those that have never used TeX or LaTeX before.  Using _R Markdown_ will also allow you to easily keep track of your analyses in **R** chunks of code, with the resulting plots and output included as well.  The hope is this _R Markdown_ template gets you in the habit of doing reproducible research, which benefits you long-term as a researcher, but also will greatly help anyone that is trying to reproduce or build onto your results down the road.

Hopefully, you won't have much of a learning period to go through and you will reap the benefits of a nicely formatted thesis.  The use of LaTeX in combination with _Markdown_ is more consistent than the output of a word processor, much less prone to corruption or crashing, and the resulting file is smaller than a Word file. While you may have never had problems using Word in the past, your thesis is likely going to be about twice as large and complex as anything you've written before, taxing Word's capabilities.  After working with _Markdown_ and **R** together for a few weeks, we are confident this will be your reporting style of choice going forward.

<!-- 
If you're still on the fence about using _R Markdown_, check out the resource
for newbies available at <https://rbasics.netlify.com>. 
-->

**Why use it?**

_R Markdown_ creates a simple and straightforward way to interface with the beauty of LaTeX.  Packages have been written in **R** to work directly with LaTeX to produce nicely formatting tables and paragraphs. In addition to creating a user friendly interface to LaTeX, _R Markdown_ also allows you to read in your data, to analyze it and to visualize it using **R** functions, and also to provide the documentation and commentary on the results of your project.  Further, it allows for **R** results to be passed inline to the commentary of your results.  You'll see more on this later.

<!-- 
Having your code and commentary all together in one place has a plethora of 
benefits!
-->

**Who should use it?**

Anyone who needs to use data analysis, math, tables, a lot of figures, complex cross-references, or who just cares about the final appearance of their document should use _R Markdown_. Of particular use should be anyone in the sciences, but the user-friendly nature of _Markdown_ and its ability to keep track of and easily include figures, automatically generate a table of contents, index, references, table of figures, etc. should make it of great benefit to nearly anyone writing a thesis project.

**For additional help with bookdown**

Please visit [the free online bookdown reference guide](https://bookdown.org/yihui/bookdown/).

>>>>

---
title: 'My Final College Paper'
author: 'Margaret A. Slein'
date: 'May 2021'
institution: 'Reed College'
division: 'Mathematics and Natural Sciences'
advisor: 'Sam Fey'
department: 'Biology'
degree: 'Bachelor of Arts'
knit: bookdown::render_book
site: bookdown::bookdown_site
params:
  'Install needed packages for {thesisdown}': True
output:
  thesisdown::thesis_pdf: default 
abstract: '`r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00-abstract.Rmd")), collapse = "\n  ")`'
acknowledgements: |

  The future is bright, the future is dark

@@

dedication: |
  You can have a dedication here if you wish. 
bibliography: bib/thesis.bib
csl: csl/apa.csl
lot: true
lof: true
---

@@

## The historical significance of variability 
It is no secret that the natural world has relied on resilience and persistence in the face of environmental change for millenia (cite). Cyanobacteria can trace their origin back to a bacteria that evolved the ability to perform oxygenic photosynthesis to persist in a changing world (Soo et al 2017, science). Similarly, vascular plants have evolved different photosynthetic systems (C3, CAM) to manage a variety of thermal and spatial conditions (cite). While cyanobacteria and vascular plants evolved new strategies and structures to persist in the environment, the environment has continued to remain dynamic, such that evolutionary responses and plasticity may become even more relevant (cite). Understanding how organisms respond to the environment and its dynamic nature is significant for how to capitulate ecosystems currently and in the future.

## Relevant significance of variability today 
In addition to the importance of environmental variability across geologic timescales, organisms also interact with environmental variability on several shorter timescales. Ecosystems and organisms inhabiting those environments have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e. ENSO). The scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems (cite). As environmental patterns continue to change, from increased mean temperatures to more severe weather events (i.e. hurricanes), understanding the limits to organisms’ ability to cope with changes in environmental conditions is increasingly important (cite).

## Why this study on variability matters 
While there is a large body of literature on environmental variability including its relevance for X, y, and Z (include citation), there is a lack of consensus as to which scales of biological organization will be most severely affected by different shifts in environmental conditions (problem statement). We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as whether different patterns of variation differentially affect different levels of organization (aims). To accomplish this, we review recent progress in the field of environmental variability, specifically changes in temperature, to identify gaps between theory, experiment, and implementation.
